%! Author = sriadityavedantam
%! Date = 3/25/26

\part{Real Analysis}

\chapter{Introduction}
\epigraph{AI is like how a chocolate croissant is... Irresistible}{\textit{Joseph H.G Fu}}

\epigraph{The natural numbers are the work of God.}{\textit{Leopold Kronecker}}

\epigraph{Sets are like machine code, its one of the ways you can have it, but you will never catch me talking about categories.}{\textit{Joseph H.G Fu}}

\epigraph{You can't get anywhere in life without desire.}{\textit{Joseph H.G Fu}}

\newpage

\epigraph{Do not have a belabored proof. If you are able to convince yourself and show you understand what is happening, then there is nothing more that is needed. Many undergraduates write too much that it makes me less convinced.}{\textit{Joseph H.G Fu}}

\newpage
\begin{itemize}
    \item First part: Undergraduate Studies. We follow Abbott Chapter 1-7.
    \item We follow Abbott `Understanding Analysis':
    \begin{itemize}
        \item Chap 1 `The Real Numbers'
        \item Chap 2 `Sequences and Series'
        \item Chap 3 `Basic Topology of $\r$'
        \item Chap 4 `Functional Limits and Continuity'
        \item Chap 5 `The Derivative'
        \item Chap 6 `Sequences and Series of Functions'
        \item Chap 7 `The Riemann Integral'
    \end{itemize}
    \item Second part: Semi-Graduate Studies. We follow Abbott Chapter 8 only.
    \item 20 graduate problems (solve them during the semester) No feedback during the semester, but asking questions is allowed. Working together is allowed.
    \item Exam (1 paged cheat-sheet, closed book, fresh dissimilar questions)
\end{itemize}

\chapter{The Real Numbers}

\input{files/Analysis/sections/day1}
\input{files/Analysis/sections/day2}
\input{files/Analysis/sections/day4}
\input{files/Analysis/sections/day5}

\chapter{Sequences and Series}

\input{files/Analysis/sections/day6}
\input{files/Analysis/sections/day7}
\input{files/Analysis/sections/day8}
\input{files/Analysis/sections/day9}
\input{files/Analysis/sections/day10}
\input{files/Analysis/sections/day12}
\input{files/Analysis/sections/day13}
\input{files/Analysis/sections/day14}

\lesson{17}{Friday 19 September 2025 9:10}{Day 17}

Test One Day.

\input{files/Analysis/sections/day19}

\chapter{Topology of $\r$}

\input{files/Analysis/sections/day21}
\input{files/Analysis/sections/day22}
\input{files/Analysis/sections/day23}
\input{files/Analysis/sections/day24}
\input{files/Analysis/sections/day25}
\input{files/Analysis/sections/day26}
\input{files/Analysis/sections/day27}

\chapter{Functional Limits and Continuity}

\input{files/Analysis/sections/day28}
\input{files/Analysis/sections/day29}
\input{files/Analysis/sections/day30}
\input{files/Analysis/sections/day31}
\input{files/Analysis/sections/day32}

\chapter{Derivatives}

\input{files/Analysis/sections/day33}
\input{files/Analysis/sections/day34}
\input{files/Analysis/sections/day35}
\input{files/Analysis/sections/day36}

\chapter{Sequences and Series of Functions}

\input{files/Analysis/sections/day38}
\input{files/Analysis/sections/day39}
\input{files/Analysis/sections/day40}

\chapter{The Riemann Integral}

\input{files/Analysis/sections/day41}
\input{files/Analysis/sections/day42}
\input{files/Analysis/sections/day43}
\input{files/Analysis/sections/day44}

\chapter{Metric Spaces}

\input{files/Analysis/sections/day45}
\input{files/Analysis/sections/day46}
\input{files/Analysis/sections/day47}
\input{files/Analysis/sections/day48}
\input{files/Analysis/sections/day49}
\input{files/Analysis/sections/day50}

\chapter{Problem Sets}

\input{files/Analysis/sections/homework0}
\input{files/Analysis/sections/homework1}
\input{files/Analysis/sections/homework2}
\input{files/Analysis/sections/homework3}
\input{files/Analysis/sections/homework4}
\input{files/Analysis/sections/homework5}