%! Author = sriadityavedantam
%! Date = 3/25/26

\section{Homework 5}

\begin{problem}[Exercise 6.2.1]{
    Let
    \[
        f_n(x)=\frac{nx}{1+nx^2}.
    \]
    \begin{enumerate}[label={(\alph*)}]
	\item Find the pointwise limit of $(f_n)$ for all $x\in(0,\infty)$.
    \item Is the convergence uniform on $(0,\infty)$?
    \item Is the convergence uniform on $(0,1)$?
    \item Is the convergence uniform on $(1,\infty)$?
    \end{enumerate}
}
    \textbf{(a).}
    For fixed $x>0$,
    \[
        f_n(x)=\frac{nx}{1+nx^2}
        =
        \frac{x}{\frac1n+x^2}
        \to \frac{1}{x}.
    \]
    So the pointwise limit is
    \[
        f(x)=\frac1x.
    \]

    \textbf{(b).}
    The convergence is not uniform on $(0,\infty)$.
    Indeed,
    \[
        \abs{f_n(x)-\frac1x}
        =
        \abs{\frac{nx}{1+nx^2}-\frac1x}
        =
        \frac{1}{x(1+nx^2)}.
    \]
    Take $x=\frac1n$.
    Then
    \[
        \abs{f_n(1/n)-n}
        =
        \frac{n^2}{n+1},
    \]
    which does not go to $0$.
    \textbf{(c).}
    The convergence is not uniform on $(0,1)$ for the same reason, since $\frac1n\in(0,1)$ for large $n$.

    \textbf{(d).}
    The convergence is uniform on $(1,\infty)$.
    If $x>1$, then
    \[
        \abs{f_n(x)-\frac1x}
        =
        \frac{1}{x(1+nx^2)}
        \leq
        \frac{1}{1+n}.
    \]
    Since $\frac{1}{n+1}\to 0$, the convergence is uniform.
\end{problem}

\begin{definition}[Uniformly Differentiable]
    A function $f$ is uniformly differentiable on $A$ if for every $\eps > 0$, there exists $\delta > 0$ such that
    \[
        0<\abs{x-y}<\delta
        \implies
        \abs{\frac{f(x)-f(y)}{x-y}-f\uptick(y)}<\eps.
    \]
\end{definition}

\begin{problem}[Exercise 6.2.4]{
    If a function is uniformly differentiable, then its derivative must be continuous.
}
    Let $\eps > 0$.
    By uniform differentiability, there exists $\delta > 0$ such that whenever
    \[
        0<\abs{x-y}<\delta,
    \]
    we have
    \[
        \abs{\frac{f(x)-f(y)}{x-y}-f'(y)}<\frac{\eps}{2}.
    \]
    Also,
    \[
        \frac{f(x)-f(y)}{x-y}
        =
        \frac{f(y)-f(x)}{y-x},
    \]
    so
    \[
        \abs{\frac{f(y)-f(x)}{y-x}-f'(x)}<\frac{\eps}{2}.
    \]
    Hence
    \begin{align*}
        \abs{f'(x)-f'(y)}
        &\leq
        \abs{f'(x)-\frac{f(y)-f(x)}{y-x}}
        +
        \abs{\frac{f(x)-f(y)}{x-y}-f'(y)}\\
        &<\frac{\eps}{2}+\frac{\eps}{2}\\
        &=\eps.
    \end{align*}
    Therefore $f'$ is continuous.
\end{problem}

\begin{problem}[Exercise 6.2.6]{
    Assume $f_n\to f$ pointwise on a set $A$.
    Provide examples showing that all of the following propositions are false if the convergence is only pointwise.
    Then decide which are true under the stronger hypothesis of uniform convergence.
    \begin{enumerate}[label={(\alph*)}]
	\item If each $f_n$ is uniformly continuous, then $f$ is uniformly continuous.
    \item If each $f_n$ is bounded, then $f$ is bounded.
    \item If each $f_n$ has a finite number of discontinuities, then $f$ has a finite number of discontinuities.
    \item If each $f_n$ has fewer than $M$ discontinuities, where $M\in\n$ is fixed, then $f$ has fewer than $M$ discontinuities.
    \item If each $f_n$ has at most a countable number of discontinuities, then $f$ has at most a countable number of discontinuities.
    \end{enumerate}
}
    \textbf{(a).}
    Take
    \[
        f_n(x)=x^n
        \qquad\text{on }[0,1].
    \]
    Each $f_n$ is uniformly continuous.
    The pointwise limit is
    \[
        f(x)=
        \begin{cases}
            0,&0\leq x<1,\\
            1,&x=1,
        \end{cases}
    \]
    which is not continuous, and therefore not uniformly continuous.
    Under uniform convergence, the statement is true.

    \textbf{(b).}
    Take
    \[
        f_n(x)=
        \begin{cases}
            x,&x<n,\\
            n,&x\geq n.
        \end{cases}
    \]
    Each $f_n$ is bounded.
    The pointwise limit is
    \[
        f(x)=x,
    \]
    which is unbounded.
    Under uniform convergence, the statement is true.

    \textbf{(c).}
    Define
    \[
        f_n(x)=
        \begin{cases}
            \lfloor x\rfloor,&x\in[0,n],\\
            x,&x\notin[0,n].
        \end{cases}
    \]
    Each $f_n$ has finitely many discontinuities.
    The pointwise limit is
    \[
        f(x)=
        \begin{cases}
            x,&x<0,\\
            \lfloor x\rfloor,&x\geq 0,
        \end{cases}
    \]
    which has infinitely many discontinuities.
    Under uniform convergence, this statement is false in general.

    \textbf{(d).}
    Take
    \[
        f_n(x)=\frac{1}{1+nx^2}.
    \]
    Each $f_n$ is continuous, so with $M=1$ each has fewer than $1$ discontinuities.
    But the pointwise limit is
    \[
        f(x)=
        \begin{cases}
            1,&x=0,\\
            0,&x\neq 0,
        \end{cases}
    \]
    which has one discontinuity.
    Under uniform convergence, the statement is true.

    \textbf{(e).}
    A standard counterexample is a sequence of functions converging pointwise to the Dirichlet function
    \[
        \chi_{\q}(x)=
        \begin{cases}
            1,&x\in\q,\\
            0,&x\notin\q.
        \end{cases}
    \]
    The limit has discontinuities at every real number, so the set of discontinuities is uncountable.
    Under uniform convergence, the statement is true.
\end{problem}

\begin{problem}[Exercise 6.2.8]{
    Let $(g_n)$ be a sequence of continuous functions that converges uniformly to $g$ on a compact set $K$.
    If $g(x)\neq 0$ on $K$, show that $(1/g_n)$ converges uniformly on $K$ to $1/g$.
}
    Since $g$ is continuous on compact $K$ and never vanishes, there exists $m>0$ such that
    \[
        \abs{g(x)}\geq m
        \qquad\text{for all }x\in K.
    \]
    Because $g_n\to g$ uniformly, choose $N_1$ such that for all $n\geq N_1$,
    \[
        \abs{g_n(x)-g(x)}<\frac{m}{2}
        \qquad\text{for all }x\in K.
    \]
    Then for $n\geq N_1$,
    \[
        \abs{g_n(x)}
        \geq
        \abs{g(x)}-\abs{g_n(x)-g(x)}
        \geq
        m-\frac{m}{2}
        =
        \frac{m}{2}.
    \]
    Now let $\eps>0$.
    Choose $N_2$ such that for all $n\geq N_2$,
    \[
        \abs{g_n(x)-g(x)}<\frac{\eps m^2}{2}
        \qquad\text{for all }x\in K.
    \]
    Then for $n\geq \max\{N_1,N_2\}$,
    \begin{align*}
        \abs{\frac1{g_n(x)}-\frac1{g(x)}}
        &=
        \abs{\frac{g(x)-g_n(x)}{g_n(x)g(x)}}\\
        &\leq
        \frac{\abs{g_n(x)-g(x)}}{(m/2)m}\\
        &<
        \frac{(\eps m^2)/2}{m^2/2}\\
        &=
        \eps.
    \end{align*}
    Hence
    \[
        \frac1{g_n}\to\frac1g
    \]
    uniformly on $K$.
\end{problem}

\begin{problem}{
    Assume $(f_n)$ and $(g_n)$ are uniformly convergent sequences of functions.
    \begin{enumerate}[label={(\alph*)}]
	\item Show that $(f_n+g_n)$ is a uniformly convergent sequence of functions.
    \item Give an example to show that the product $(f_ng_n)$ may not converge uniformly.
    \item Prove that if there exists an $M>0$ such that $\abs{f_n}\leq M$ and $\abs{g_n}\leq M$ for all $n\in\n$, then $(f_ng_n)$ does converge uniformly.
    \end{enumerate}
}
    \textbf{(a).}
    Suppose
    \[
        f_n\to f
        \qquad\text{and}\qquad
        g_n\to g
    \]
    uniformly.
    Let $\eps>0$.
    Choose $N_1$ such that for all $n\geq N_1$,
    \[
        \abs{f_n(x)-f(x)}<\frac{\eps}{2}
        \qquad\text{for all }x.
    \]
    Choose $N_2$ such that for all $n\geq N_2$,
    \[
        \abs{g_n(x)-g(x)}<\frac{\eps}{2}
        \qquad\text{for all }x.
    \]
    Then for all $n\geq \max\{N_1,N_2\}$,
    \begin{align*}
        \abs{(f_n+g_n)(x)-(f+g)(x)}
        &\leq
        \abs{f_n(x)-f(x)}+\abs{g_n(x)-g(x)}\\
        &<\eps.
    \end{align*}
    So $f_n+g_n\to f+g$ uniformly.
    \textbf{(b).}
    Take
    \[
        f_n(x)=x+\frac1n,
        \qquad
        g_n(x)=x+\frac1n
        \qquad\text{on }\r.
    \]
    Then $f_n\to x$ and $g_n\to x$ uniformly on $\r$.
    But
    \[
        f_n(x)g_n(x)-x^2
        =
        \frac{2x}{n}+\frac1{n^2},
    \]
    and the supremum of this difference over $\r$ is not finite.
    So $(f_ng_n)$ does not converge uniformly on $\r$.
    \textbf{(c).}
    Assume
    \[
        f_n\to f,
        \qquad
        g_n\to g
    \]
    uniformly, and
    \[
        \abs{f_n(x)}\leq M,
        \qquad
        \abs{g_n(x)}\leq M
    \]
    for all $x$ and all $n$.
    Then also
    \[
        \abs{f(x)}\leq M,
        \qquad
        \abs{g(x)}\leq M.
    \]
    Now
    \begin{align*}
        \abs{f_n(x)g_n(x)-f(x)g(x)}
        &\leq
        \abs{f_n(x)}\abs{g_n(x)-g(x)}
        +
        \abs{g(x)}\abs{f_n(x)-f(x)}\\
        &\leq
        M\abs{g_n(x)-g(x)}
        +
        M\abs{f_n(x)-f(x)}.
    \end{align*}
    Let $\eps>0$.
    Choose $N$ such that for all $n\geq N$,
    \[
        \abs{f_n(x)-f(x)}<\frac{\eps}{2M}
        \qquad\text{and}\qquad
        \abs{g_n(x)-g(x)}<\frac{\eps}{2M}
    \]
    for all $x$.
    Then for $n\geq N$,
    \[
        \abs{f_n(x)g_n(x)-f(x)g(x)}<\eps
    \]
    for all $x$.
    Hence $(f_ng_n)$ converges uniformly to $fg$.
\end{problem}